---
title: "19-3"
author: "Nikhil Gopal"
date: "3/14/2022"
output: pdf_document
---

```{r, message=FALSE, echo=FALSE}
rm(list = ls())
```


**19.4**
*Pre-test and post-test:* 100 students are given a pre-test, then a treatment or control is randomly assigned to each, then they get a post-test. Given the following regression model:

$$post-test = \alpha + b * pre-test + \theta * z + error$$  

where z = 1 for treated units and theta for controls. Further suppose that pre_test has mean 40 and standard deviation 15. Suppose b = 0.7 and theta = 10 and the mean for post_test is 50 for the students in the control group. Further suppose that the residual standard deviation of the regression is 10.

+ a) Determine *a*


The mean of the post_test score is 50, and the mean of the pre_treatment score is 40, so the effect size $\alpha = 50-40 = 10$

+ b) What is the standard deviation of the post-test scores for the students in the control group?

It should be 10 + the error.

+ c) What are the mean and standard deviation of the post-test scores in the treatment group?

$$posttest = 10 + 0.7(40) + 10(1) + error$$

$$posttest = 48$$

The mean will be 48 and the standard deviation will be the error.
**19.5**

*Causal inference using logistic regression*: Suppose you have fit a model:


```{r, eval=FALSE}
fit <- stan_glm(y ~ z + age + z:age, family=binomial(link="logit"), data=mydata)
```


with binary outcome y, treatment indicator z, and age measured in years. Give R code to
produce an estimate and standard error of average treatment effect in a large population, given a
vector n_pop of length 82 that has the number of people in the population at each age from 18
through 99:

```{r, eval=FALSE}
predictions$effect <- predict(fit, n_pop)

avg_effect <- mean(predictions$effect)

st_err <-sd(predictions$effect)/length(n_pop)
```

