---
title: "19-3"
author: "Nikhil Gopal"
date: "3/14/2022"
output: pdf_document
---

```{r, message=FALSE, echo=FALSE}
rm(list = ls())
library(rstanarm)
library(rosdata)
library(tidyverse)
```


**19.4**
*Pre-test and post-test:* 100 students are given a pre-test, then a treatment or control is randomly
assigned to each, then they get a post-test. Given the following regression model:

                            post_test = a + b * pre_test + theta * z + error  

where z = 1 for treated units and theta for controls. Further suppose that pre_test has mean 40
and standard deviation 15. Suppose b = 0.7 and theta = 10 and the mean for post_test is 50 for
the students in the control group. Further suppose that the residual standard deviation of the
regression is 10.

+ a) Determine *a*


+ b) What is the standard deviation of the post-test scores for the students in the control group?
+ c) What are the mean and standard deviation of the post-test scores in the treatment group?


**19.5**

*Causal inference using logistic regression*: Suppose you have fit a model:


```{r, eval=FALSE}
fit <- stan_glm(y ~ z + age + z:age, family=binomial(link="logit"), data=mydata)
```


with binary outcome y, treatment indicator z, and age measured in years. Give R code to
produce an estimate and standard error of average treatment effect in a large population, given a
vector n_pop of length 82 that has the number of people in the population at each age from 18
through 99:

```{r, eval=FALSE}

predictions$effect <- predict(fit, n_pop)

avg_effect <- mean(exp(predictions$effect))

st_err <-sd(predictions$effect)/sqrt(82)
```

