---
title: "Untitled"
author: "Nikhil Gopal"
date: "2/5/2022"
output: html_document
---

**13.13 partner project**

```{r, message=FALSE}
rm(list = ls())
library(tidyverse)
library(rstanarm)
```

In this project, we will seek to fit a logistic regression model to predict a state's outcome in the 2020 US presidential election, with income and percentage of white residents as predictors.

**Import and Clean Data**

Presidential election outcome data source:

https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/42MVDX

Income data source:

https://www.kff.org/other/state-indicator/median-annual-income/?currentTimeframe=0&sortModel=%7B%22colId%22:%22Location%22,%22sort%22:%22asc%22%7D

Race data source:

https://www.kff.org/other/state-indicator/population-distribution-by-race-ethnicity-cps/?currentTimeframe=0&sortModel=%7B%22colId%22:%22Location%22,%22sort%22:%22asc%22%7D


```{r}

presidental <- read.csv("2020outcomes.csv")
presidental <- presidental[1:2]

#note that you have to delete the first 2 rows to get the data frame to be recognized by R
income <- read.csv("income.csv")
income <- income[2:52,1:2]
colnames(income) <- c("state", "income")

#note that you have to delete the first 2 rows to get the data frame to be recognized by R
race <- read.csv("race.csv")
race <- race[2:52,1:2]
colnames(race) <- c("state", "pct_white")

df <- merge(income,race,by = "state")
df <- merge(df, presidental, by = "state")
```


**Fit Logistic Regression Model:**

```{r, message=FALSE}
#recode D/R to 1/0
df$called <- recode(df$called, D = 1, R = 0)

df$income <- gsub("\\$", "", df$income)
df$income <- gsub("\\,", "", df$income)
df$income <- as.numeric(df$income)

fit <- stan_glm(called ~ income + pct_white, family = binomial(link = "logit"), data = df)

fit2 <- stan_glm(called ~ pct_white, family = binomial(link = "logit"), data = df)
```

Print the summary:

```{r}
print(fit)
```

Graph the model (remove income to make it easier to graph as coefficient was 0):

```{r}

plot(df$pct_white, invlogit(coef(fit2)[1] + coef(fit2)[2]*df$pct_white))
curve(invlogit(coef(fit2)[1] + coef(fit2)[2]*x), xlim = c(0, 1), from = 0, to = 1, col = "red", add = T)
```

We fit a model predicting weather or not a state would vote for Democrats in the 2020 presidential election. We used the percentage of white residents in the state as a predictor and  removed the income variable from the model, as the coefficient was zero when we originally fit. This coefficient value predicts that the log odds of a state voting democrat decreases by 3.8 for every 1 percentage increase in a state's white population. This effect is also shown in the graph above.


