---
title: '14.3'
author: "Nikhil Gopal"
date: "2/8/2022"
output:
  html_document:
    df_print: paged
---

```{r, message=FALSE}
rm(list = ls())
library(rstanarm)
library(tidyverse)
```

**14.3**

Graphing logistic regressions: The well-switching data described in Section 13.7 are in the
folder Arsenic

A:

Fit a logistic regression for the probability of switching using log (distance to nearest safe
well) as a predictor.

```{r}
wells$dist100 <- wells$dist/100

fit_1 <- stan_glm(formula = switch ~ log(dist), family=binomial(link="logit"),
          data=wells)


print(fit_1)
```

B:

Make a graph similar to Figure 13.8b displaying Pr(switch) as a function of distance to
nearest safe well, along with the data.

```{r}
#code from p 252 in text
jitter_binary <- function(a, jitt=0.05){
ifelse(a==0, runif(length(a), 0, jitt), runif(length(a), 1 - jitt, 1))
}

wells$switch_jitter <- jitter_binary(wells$switch)
plot(wells$dist, wells$switch_jitter, xlab = "Distance from Well", ylab = "Pr(Switch)")
curve(invlogit(coef(fit_1)[1] + coef(fit_1)[2]*x), add=TRUE)

```

C:

Make a residual plot and binned residual plot as in Figure 14.8.

```{r}
#Residual plot
plot(fit_1$fitted.values, fit_1$residuals, main = "Residual Plot",
     xlab = "Estimated Pr(switching)", ylab = "Residual")


#using rstanarm library
pp_check(fit_1, plotfun = "error_binned")


#using arm (https://rdrr.io/cran/arm/man/binnedplot.html), more like book graph
arm::binnedplot(x = fit_1$fitted.values, fit_1$residuals,
                xlab = "Estimated Pr(Switching)", 
                main = "Binned Residual vs. Predicted Values", 
                col.int = "gray", nclass = 40)
```

D:

Compute the error rate of the fitted model and compare to the error rate of the null model.


```{r}
#calculate predictions
wells$predicted <- predict(fit_1, newdata = wells, type = "response")

#fit null model
null_model <- stan_glm(formula = switch ~ 1, family=binomial(link="logit"),
          data=wells)

#code from p 273 to compute error rate of models
error_rate_fit <- mean((wells$predicted>0.5 & wells$switch==0) | (wells$predicted<0.5 & wells$switch==1))

#We can compare it to the error rate of the null model, which is simply to assign the same probability to each yi. This is simply logistic regression with only a constant term, and the estimated probability will simply be the proportion of 1’s in the data, or p = n i=1 yi/n (recalling that each yi = 0 or 1). The error rate of the null model is then p or 1−p, whichever is lower

prop_1_in_data <- table(wells$switch)

prop_1_in_data <- 1737/3020

error_rate_null <- 1 - prop_1_in_data

df <- data.frame(
  "model" <- c("Fitted", "Null"),
  "error" <- c(error_rate_fit, error_rate_null))

colnames(df) <- c("model", "error")

print(paste("Fitted model error: ", round(error_rate_fit, 3)))
print(paste("Null model error: ", round(error_rate_null, 3)))

bp <- barplot(df$error, names.arg = df$model, main = "Barplot of Fitted vs Null Model Error")
text(bp, 0, round(df$error, 3),cex=1,pos=3) 
```

E:

Create indicator variables corresponding to dist < 100; dist between 100 and 200; and
dist > 200. Fit a logistic regression for Pr(switch) using these indicators. With this new
model, repeat the computations and graphs for part (a) of this exercise.

Create indicator variables and fit model:

```{r}

wells$less_100 <- ifelse(wells$dist < 100, 1, 0)
wells$greater_200 <- ifelse(wells$dist > 200, 1, 0)
wells$between <- ifelse(wells$dist < 200 & wells$dist > 100, 1, 0)

wells$less_100 <- as.factor(wells$less_100)
wells$greater_200 <- as.factor(wells$greater_200)
wells$between <- as.factor(wells$between)

fit_2 <- stan_glm(formula = switch ~ less_100 + greater_200 + between, family=binomial(link="logit"),
          data=wells)

```

Distance from well vs Pr(switch):
```{r}
#code from p 252 in text
jitter_binary <- function(a, jitt=0.05){
ifelse(a==0, runif(length(a), 0, jitt), runif(length(a), 1 - jitt, 1))
}

wells$switch_jitter <- jitter_binary(wells$switch)
plot(wells$dist, wells$switch_jitter, xlab = "Distance from Well", ylab = "Pr(Switch)")
curve(invlogit(coef(fit_2)[1] + coef(fit_2)[2]*x), add=TRUE)

```

Residual and Binned Plots:
```{r}
#Residual plot
plot(fit_2$fitted.values, fit_2$residuals, main = "Residual Plot",
     xlab = "Estimated Pr(switching)", ylab = "Residual")


#using rstanarm library
pp_check(fit_2, plotfun = "error_binned")


#using arm (https://rdrr.io/cran/arm/man/binnedplot.html), more like book graph
arm::binnedplot(x = fit_2$fitted.values, fit_2$residuals,
                xlab = "Estimated Pr(Switching)", 
                main = "Binned Residual vs. Predicted Values", 
                col.int = "gray", nclass = 40)
```


Error rate calculation:

```{r}
#calculate predictions
wells$predicted_2 <- predict(fit_2, newdata = wells, type = "response")


#code from p 273 to compute error rate of models
error_rate_fit_2 <- mean((wells$predicted_2>0.5 & wells$switch==0) | (wells$predicted_2<0.5 & wells$switch==1))

#We can compare it to the error rate of the null model, which is simply to assign the same probability to each yi. This is simply logistic regression with only a constant term, and the estimated probability will simply be the proportion of 1’s in the data, or p = n i=1 yi/n (recalling that each yi = 0 or 1). The error rate of the null model is then p or 1−p, whichever is lower

prop_1_in_data <- table(wells$switch)

prop_1_in_data <- 1737/3020

error_rate_null <- 1 - prop_1_in_data

df2 <- data.frame(
  "model" <- c("Fitted", "Null"),
  "error" <- c(error_rate_fit_2, error_rate_null))

colnames(df2) <- c("model", "error")

print(paste("Fitted model error: ", round(error_rate_fit_2, 3)))
print(paste("Null model error: ", round(error_rate_null, 3)))

bp <- barplot(df$error, names.arg = df2$model, main = "Barplot of Fitted vs Null Model Error")
text(bp, 0, round(df2$error, 3),cex=1,pos=3) 
```

This model had an error rate of 0.4092715, which is slightly higher compared to our original model's error rate of 0.4056291. Thus, in terms of making predictions on this dataset, our original model is better than using this indicator model.
